---
title: " Practical Machine Learning - Course Project: Writeup"
author: "Cleber R. Guirelli"
output: html_document
---
 
### Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

### Data 


The training data for this project are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment. 

### Loading Data

First we load necessary libraries:

```{r}
suppressMessages(library(ggplot2))
suppressMessages(library(lattice))
suppressMessages(library(caret))
suppressMessages(library(randomForest))
suppressMessages(library(e1071))
suppressMessages(library(gbm))
suppressMessages(library(doParallel))
suppressMessages(library(plyr))
```
Then the data:

```{r}
train_data =  read.csv("pml-training.csv", na.strings=c("#DIV/0!"), row.names = 1)
test_data =  read.csv("pml-testing.csv", na.strings=c("#DIV/0!"), row.names = 1)
dim(train_data)
table(train_data$classe)
```

### Cleaning Data

Using *View(train_data)* we see that the first 6 columns have no use in prediction and we have a lot of NA values. So let's clean data a bit removing columns with more than 90% of NA or "" :

```{r}
train_data <- train_data[, 6:dim(train_data)[2]]
test_data <- test_data[, 6:dim(test_data)[2]]

limit <- dim(train_data)[1] * 0.9

good_data <- !apply(train_data, 2, 
                     function(x) sum(is.na(x)) > limit  || sum(x=="") > limit)

train_data <- train_data[, good_data]
test_data <- test_data[, good_data]
bad_data <- nearZeroVar(train_data, saveMetrics = TRUE)
train_data <- train_data[, bad_data$nzv==FALSE]
test_data <- test_data[, bad_data$nzv==FALSE]

```

### Creating Model and Validating

Partition of data into training and test model:

```{r}
set.seed(1970)
train_data$classe = factor(train_data$classe)
train_set <- createDataPartition(train_data$classe, p = 0.8, list = FALSE)
train_model_data <- train_data[train_set, ]
test_model_data <- train_data[-train_set, ]
```

Create 2 different models and test the models with predicitons. 
odel *rf* had 0.999 accuracy with test datset but takes a very long time to run.
Model *gbm* has 0.9921 and runs faster, so it was used for prediction.

```{r}
# #Random forests - Accuracy : 0.999  
# model_1 <- train(classe ~ ., data=train_model_data, method="rf")
# prediction_1 <- predict(model_1, test_model_data)
# confusionMatrix(prediction_1, test_model_data$classe)
# model_1$times
# predict_submission = predict(model_1, test_data)

#Boosted trees - Accuracy : 0.9921
model_2 <- train(classe ~ ., data=train_model_data, method="gbm",verbose = FALSE)
prediction_2 <- predict(model_2, test_model_data)
confusionMatrix(prediction_2, test_model_data$classe)
```

Using Predicted Model, generate answer for test data:

```{r}
predict_submission = predict(model_2, test_data)

pml_write_files = function(x){
    n = length(x)
    for(i in 1:n){
        filename = paste0("problem_id_",i,".txt")
        write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
    }
}
pml_write_files(predict_submission)
```
All 20 answers were correct.
